---
title: "Simple R Functions "
author: "Natalya Shelchkova"
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$

```{r}

tmpFn3 <- function(x, n){
  xVec <- rep(x,n)
  return (1 + sum((xVec^(1:n))/(1:n)))
}

tmpFn3(2, 3)

```

\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))

```{r}

tmpFn4 <- function(xVec){
  n <- length(xVec)
  return((xVec[1:(n-2)]+xVec[2:(n-1)] + xVec[3:n])/3)
}
tmpFn4(c(1:5,6:1))

```

### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.

```{r}

tmpFn5 <- function(xVec){
  return(ifelse(xVec < 0, xVec^2+2*xVec+3, ifelse(xVec < 2, xVec+3, xVec^2+4*xVec-7)))
}

plot(tmpFn5(-3:3))

```


### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$

```{r}
tmpFn6 <- function(A){
  return(ifelse(A%%2 == 1, A*2, A))
}

tmpFn6(matrix(c(1,1,3,5,2,6,-2,-1,-3), byrow = TRUE, nr = 3))

```
### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$

```{r}

tmpFn7 <- function(n, k){
  return(diag(k, nr = n))
}

```
### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.

```{r}

quadrant <- function(alpha){
  if (alpha > 360){
    alpha <- alpha%%360
  }
  if (alpha >= 0 & alpha < 90){
    print("Quadrant 1")
  }
  else if (alpha >= 90 & alpha < 180){
    print("Quadrant 2")
  }
  else if (alpha >= 180 & alpha < 270){
    print("Quadrant 3")
  }
  else {
    print("Quadrant 4")
  }
}

quadrant(90)

```

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

```{r}

weekday <- function(day, month, year){
  month_vector <- c(11:12,1:10)
  m <- month_vector[month[1]]
  k <- day[1]
  tmp_year <- as.numeric(strsplit(as.character(year[1]), "")[[1]])
  mod_year <- c(as.numeric(paste0(tmp_year[1],tmp_year[2])), as.numeric(paste0(tmp_year[3],tmp_year[4])))
  if (m == 11 || m == 12){
    y <- (mod_year[2] - 1)
    if (y == 99){
      c <- (mod_year[1] - 1)
    }
    else{
      c <- mod_year[1]
    }
  } 
  else {
    c <- mod_year[1]
    y <- mod_year[2] 
  }
  f <- (floor((2.6*m) - 0.2) + k + y + floor(y/4) + floor(c/4))%%7
  return (f)
  
}

```


(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?

Yes, depending on whether day, month, and year are individual vectors or if its one vector that has [day, month, year].

### 8.

(a) Suppose $x_0 = 1$ and $x_1 = 2$ and $$x_j = x_{j-1} + \frac{2}{x_{j-1}} \textit{  for } j = 1,2,\ldots$$

Write a function \texttt{testLoop} which takes the single argument \textit{n} and returns the first \textit{n}-1 values of the sequence ${x_j}_{j \geq 0}:$ that means the values of $x_0,x_1,x_2,\ldots,x_{n-2}$.

```{r}

testLoop <- function(n){
  x <- 1:(n-1)
  for (ii in 3:(n-1)){
    x[ii] <- x[ii-1] + (2/x[ii-1])
  }
  return(x)
}

```

(b) Now write a function \texttt{testLoop2} which takes a single argument \texttt{yVec} which is a vector. The function should return

$$ \sum_{j=1}^n e^j $$ where \textit{n} is the length of \texttt{yVec}.

```{r}

testLoop2 <- function(yVec){
  n <- length(yVec)
  return(sum(exp(yVec)^(1:n)))
}

```

### 9.

(a) Write a function \texttt{quadmap(start, rho, niter)} which returns the vector $(x_1,\ldots,x_n)$ where $x_k = rx_{k-1}(1-x_{k-1})$ and

\texttt{niter} denotes $n$

\texttt{start} denotes $x_1$, and

\texttt{rho} denotes $r$.

Try out the function you have written:

$\bullet$ for $r=2$ and $0<x_1<1$ you should get $x_n \rightarrow 0.5$ as $n\rightarrow \infty$.

$\bullet$ try \texttt{tmp <- quadmap(start=0.95, rho=2.99, niter = 500)}

```{r}

quadmap <- function(start, rho, niter){
  x <- 1:niter
  x[1] <- start
  for (ii in 2:niter){
    x[ii] <- rho*x[ii-1]*(1-x[ii-1])
  }
  return(x)
}

tmp <- quadmap(0.95, 2.99, 500)
```

Now switch back to the Commands window and type:

\texttt{plot(tmp, type ="l")}

Also try the plot \texttt{plot(tmp[300:500], type="l")}

(b) Now write a function which determines the number of iterations needed to get $|x_n - x_{n-1}| < 0.02$. So this function has only 2 arguments: \texttt{start} and \texttt{rho}. (For \texttt{start = 0.95} and \texttt{rho=2.99}, the answer is 84.)

```{r}

number_of_iterations <- function(start, rho){
  count <- 0
  diff <- 100000
  x_n <- start
  while (diff >= 0.02){
    x_n1 <- rho*x_n*(1-x_n)
    diff <- abs(x_n1 - x_n)
    count <- count + 1
    x_n <- x_n1
  }
  return(count)
}

```
### 10.

(a) Given a vector $(x-1,\ldots,x_n)$ the sample autocorrelation of lag $k$ is defined to be
$$ r_k = \frac{\Sigma_{i=k+1}^n (x_i - \overline{x})(x_{i-k} - \overline{x})}{\Sigma_{i=1}^n (x_i - \overline{x})^2} $$
Thus 
$$ r_1 = \frac{\Sigma_{i=2}^n (x_i - \overline{x})(x_{i-k} - \overline{x})}{\Sigma_{i=1}^n (x_i - \overline{x})^2} = \frac{(x_2 - \overline{x})(x_1 - \overline{x})+\dots+(n_n-\overline{x})(x_{n-1}-\overline{x})}{\Sigma_{i=1}^n (x_i - \overline{x})^2} $$
Write a function \texttt{tmpFn(xVec)} which takes a single argument \texttt{xVec} which is a vector and returns a \texttt{list} of two values: $r_1$ and $r_2$.
In particular, find $r_1$ and $r_2$ for the vector (2,5,8,\ldots,53,56).

```{r}

tmpFn8 <- function(xVec){
  x_mean <- mean(xVec)
  n <- length(xVec)
  r <- 1:length(xVec)
  for (k in 1:(length(xVec) - 1)){
    i <- k+1
    r[k] <- sum((xVec[i:n] - x_mean) *(xVec[k:(n-1)] - x_mean))/sum((xVec[k:n] - x_mean)^2)
  }
  return(r)
}

tmpFn8(seq(2,56,3))[1:2]
```

(b) (Harder.) Generalise the function so that it takes two arguments: the vector \texttt{xVec} and an integer \texttt{k} which lies between 1 and $n-1$ where $n$ is the length of \texttt{xVec}

The function should return a vector of the values $(r_0 = 1,r_1,\ldots,r_k)$.
If you used a loop to answer part (b), then you need to be aware that much, much better solutions are possible _ see exercises 4. (Hint: \texttt{sapply.})


```{r}

tmpFn9 <- function(xVec, k){
  x_mean <- mean(xVec)
  n <- length(xVec)
  r <- 1:(k)
  for (j in 1:(k)){
    i <- j+1
    r[j] <- sum((xVec[i:n] - x_mean) *(xVec[j:(n-1)] - x_mean))/sum((xVec[j:n] - x_mean)^2)
  }
  return(append(1,r))
}

tmpFn9(seq(2,56,3), 5)


```

